---
title: "Comparision of Covid-19 confirmed and death number between the US and China"
author: "Jiawen Chen"
date: "2022-12-06"
output:
  pdf_document:
    toc: yes
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
if(!require(DT)) install.packages("DT")
library(DT)
library(knitr)
library(readr)
library(readxl)
library(lubridate)
library(psych)
library(ggplot2)
library(data.table)
library(dplyr)
library(leaflet)
library(tidyr)
library(tidyselect)
if(!require(ggformula)) install.packages("ggformula")
if(!require(ggstance)) install.packages("ggstance")
if(!require(gghighlight))install.packages("gghighlight")
library(gghighlight)
library(ggformula)
library(ggstance)
```
# 1.Backgroud story & Introduction


I was in the US at the beginning of 2020, and then back to China in March 2020 and quarentined. So I 
observed how did the pandemic spreaded so quickly and sharply increased around the world and specificly in the US. On the other side, after the few first months in 2020, the pandemic was considered as well controlled. Therefore, in this project, firstly I will briefly check through the overall COVID-19 situation in the world wide. And then I will compare with those two countries in terms of confirmed and deaths number.

# 2.Methods:(include how and where the data were acquired, how you cleaned and wrangled the data, what tools you used for data exploration)

## About the data 

I downloaded the datasets from "https://data.humdata.org/dataset/novel-coronavirus-2019-ncov-cases?"in CSV format and updated daily and then chose the global confirmed and deaths throughout the whole pandemic (1/22/20-recent two days). The datas are updated everyday by John Hopkins center.  

## Preparation

However, the format of two datasets are about the accumulated-confirmed and death cases of different countries. So I did some filter out steps of countries/regions, cleaned and normalized that data, for example tidying dates and consolidating them into normalized time series. We have variables called "province" ""Country/Region", "dates" and Lat" and "Long" representing different regions. Also I select and transformed these two countries from the total death dataset into dataframes

```{r message=FALSE, echo=FALSE, warning=FALSE}
download.file("https://data.humdata.org/hxlproxy/api/data-preview.csv?url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_confirmed_global.csv&filename=time_series_covid19_confirmed_global.csv", "time_series_covid19_confirmed_global.csv", method="libcurl", timeout = 60)
newconfirmed <- read.csv("time_series_covid19_confirmed_global.csv", header = TRUE, na.strings = c("", " "))
 
download.file("https://data.humdata.org/hxlproxy/api/data-preview.csv?url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_deaths_global.csv&filename=time_series_covid19_deaths_global.csv", "time_series_covid19_deaths_global.csv", methos="libcurl", timeout = 60)
deaths <- read.csv("time_series_covid19_deaths_global.csv", header = TRUE, na.strings = c("", " "))
```


# 3. Here I created a heat map to shows how the confirmed cases spread in the world wide
```{r message=FALSE, echo=FALSE, warning=FALSE}
nc <- ncol(newconfirmed)
totalconfirmed <- newconfirmed[,c(1:4,nc)]
colnames(totalconfirmed)[5] = "sumconfirmed"
head(totalconfirmed)

countrytotalconfirmed <- totalconfirmed %>%
  group_by(Country.Region) %>%
  summarize(sumconfirmed=sum(sumconfirmed),
            Lat=mean(Lat, na.rm=T),
            Long=mean(Long, na.rm=T))

#total for the whole country 
confirmed.pal <- colorNumeric(c('darkgreen','goldenrod','brown'),domain=countrytotalconfirmed$sumconfirmed)
confirmedmap <- leaflet(countrytotalconfirmed) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~Lat, lng=~Long,
    label = ~paste0(sumconfirmed), color = ~ confirmed.pal(sumconfirmed),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=confirmed.pal, values=countrytotalconfirmed$sumconfirmed,
          title='Confirmed number', opacity=0.5)
confirmedmap
```

From the map, the legend shows the range of confirmed number. The red color means higher confirmed number. We can see that there America has the red dot, then India, Brazil and Germany have yellow dots. However, one drawback of this dataset is that some country, for example China's data is updated by Province, but the data from the US as a country total.

# 4. Table to check total confirmed cases for each country 
```{r message=FALSE, echo=FALSE, warning=FALSE}
datatable(totalconfirmed)
```
We can also check and sort which are the top countries have the greatest or smallest cases number. Corresponding with the previous heat map, the top five countries have the greatest cases number are US, India and France and Germany 

# 5. Transform the total death dataset, confirmed and deaths data of these two countries into dataframes 
```{r message=FALSE, echo=FALSE, warning=FALSE}
load("confirmed.Rdata")
chinaconfirmed <-
  confirmed %>%
  filter(Country.Region == "China", Province.State !="Unknown")

usconfirmed <-
  confirmed %>%
  filter(Country.Region == "US")

load("deaths.Rdata")
chinadeaths <-
  deaths %>%
  filter(Country.Region == "China", Province.State !="Unknown")

usdeaths <-
  deaths %>%
  select(.,-Province.State) %>%
  filter(Country.Region == "US")
```

# 6.graphs of each province in china and US confirmed and deaths cases. 
We only have the whole US data (because the US only count its total number?), so I just present the US as a country total.China has 50+ countries in total, so I choose the top 7 provinces that have the largest number of deaths of deaths cases. For the death cases, I only choose the provinces have more than 1000 death cases.
```{r message=FALSE, echo=FALSE, warning=FALSE}
load("confirmed.Rdata")
g1 <-
chinaconfirmed %>%
  group_by(Date) %>%
  ggplot(aes(x = Date, y = Confirmed, color=Province.State, line=Province.State)) +
  ylab("The China Confirmed Number")+
  gghighlight(max(Confirmed) > 10000)+
    geom_point(size=0.1, alpha=1)
ggplotly(g1)

load("confirmed.Rdata")
g2 <-
usconfirmed %>%
  group_by(Date) %>%
  ggplot(aes(x = Date, y = Confirmed)) + 
  ylab("The US Confirmed Number")+
  geom_point(size=0.5, alpha=1) 
ggplotly(g2)

g3 <-
chinadeaths %>%
  group_by(Date) %>%
  ggplot(aes(x = Date, y = Confirmed, color=Province.State, line=Province.State)) + 
  ylab("China Deaths Number")+
  gghighlight(max(Confirmed) > 1000)+
  geom_point(size=0.5, alpha=1) 
ggplotly(g3)

g4 <-
usdeaths %>%
  group_by(Date) %>%
  ggplot(aes(x = Date, y = Confirmed)) + 
  ylab("The US Deaths Number") +
  geom_point(size=0.5, alpha=1) 
ggplotly(g4)
```
Hongkong is the city with the most COVID-19 confirmed cases according to the graphs, also the only one with more than 5000 death cases. 

# 7. let's make those two contries in comparision
```{r message=FALSE, echo=FALSE, warning=FALSE}
  #Confirmed case
  g5 <- 
  confirmed %>%
  filter(Country.Region %in% c("US", "China")) %>%
  group_by(Country.Region, Date) %>% 
  summarise(Confirmed = sum(Confirmed)) %>% 
  gf_point(Confirmed ~ Date) %>%
  gf_facet_grid(Country.Region ~ ., scales = "free")+ylab("Confirmed case number")
ggplotly(g5)
 
  #Deaths case
g6 <-
  deaths %>%
  filter(Country.Region %in% c("US", "China")) %>%
  group_by(Country.Region, Date) %>% 
  summarise(Confirmed = sum(Confirmed)) %>% 
  gf_point(Confirmed ~ Date) %>%
  gf_facet_grid(Country.Region ~ ., scales = "free") + ylab("Deaths case number") 
ggplotly(g6)
```
We can see China started confirming cases ealier than the US but the confirmed and deaths cases increased slightly until the beginning of 2022. However, the overall confirmed and deaths cases in the US (Until 11/29/2022, deaths cases is 1.09M, from Wikipedia) are significantly higher than China (until 11/29/2022 deaths cases is 5233, from Wikipedia). Part of the reason is probably because of the strict COVID-19 policies. 